Model definition

### 2025-05-24 nbs/tutorials/models.qmd
docs: update models tutorial to exclusively use create_legacy_model2
- Replace create_standard_model with create_legacy_model2 throughout tutorial
- Update all model names and velocity keys to use 'legacy2' naming
- Emphasize that create_legacy_model2 (add_offset=True) replicates original PyroVelocity
- Update visualization section to show legacy model results
- Maintain consistency with legacy model focus across all examples
- Update uncertainty field references to match legacy2 model naming

```python
import torch
import pyro
import numpy as np
import matplotlib.pyplot as plt
import scanpy as sc
import scvelo as scv
from pyrovelocity.utils import print_anndata, anndata_string, print_string_diff
from pyrovelocity.models.modular import PyroVelocityModel
from pyrovelocity.models.modular.factory import create_standard_model
from pyrovelocity.io.datasets import bifurcation_14

@@

# Create a standard model with default components
model = create_standard_model()
print(model)

@@

# Generate posterior samples and compute velocity in one step
# The new unified get_velocity method computes both velocity and uncertainty
adata_with_velocity = model.get_velocity(
    adata=adata,
    num_samples=100,  # Use more samples for better uncertainty estimates
    basis='umap',
    compute_uncertainty=True,
    uncertainty_method='std',
    model_name='pyrovelocity'
)

@@

# Train the model
custom_model.train(
    adata=adata_with_velocity,  # Use the data with embeddings
    max_epochs=50,
    learning_rate=0.01,
    use_gpu=False
)

# Compute velocity with the custom model
custom_adata = custom_model.get_velocity(
    adata=adata_with_velocity,
    num_samples=100,
    basis='umap',
    compute_uncertainty=True,
    model_name='custom_legacy'
)
```

@@

```python
from pyrovelocity.models.modular.factory import (
    create_standard_model,
    create_model_from_config,
    create_legacy_model1,
    create_legacy_model2
)
from pyrovelocity.models.modular.config import ModelConfig

# Create a standard model (recommended for new projects)
standard_model = create_standard_model()

# Create a model that replicates the legacy implementation (add_offset=False)
legacy_model1 = create_legacy_model1()

# Create a model that replicates the legacy implementation (add_offset=True)
legacy_model2 = create_legacy_model2()

# Create a model from a configuration
config = ModelConfig.standard()
config_model = create_model_from_config(config)

# Train and use any of these models
legacy_model1.train(adata=adata_with_velocity, max_epochs=50)

# The new unified get_velocity method returns the mutated AnnData object
adata_legacy = legacy_model1.get_velocity(
    adata=adata_with_velocity,
    num_samples=100,
    basis='umap',
    compute_uncertainty=True,
    model_name='legacy1'
)

print("Available velocity layers:")
for layer in adata_legacy.layers.keys():
    if 'velocity' in layer:
        print(f"- {layer}")
```

@@

# Plot 1: Standard velocity embedding
scv.pl.velocity_embedding_grid(
    adata_with_velocity,
    basis='umap',
    color='leiden',
    title='PyroVelocity Standard Model',
    vkey='velocity_pyrovelocity',  # Use the velocity key added by get_velocity
    s=15,
    alpha=1,
    linewidth=0.5,
    ax=axes[0],
    show=False,
    legend_loc='right margin',
    density=0.4,
    scale=0.2,
    arrow_size=2,
    arrow_length=2,
    arrow_color='black',
)

# Plot 2: Legacy model comparison (if computed)
if 'velocity_legacy1' in adata_legacy.layers:
    scv.pl.velocity_embedding_grid(
        adata_legacy,
        basis='umap',
        color='leiden',
        title='PyroVelocity Legacy Model',
        vkey='velocity_legacy1',
        s=15,
        alpha=1,
        linewidth=0.5,
        ax=axes[1],
        show=False,
        legend_loc='right margin',
        density=0.4,
        scale=0.2,
        arrow_size=2,
        arrow_length=2,
        arrow_color='black',
    )
else:
    axes[1].text(0.5, 0.5, 'Legacy model\nnot computed',
                ha='center', va='center', transform=axes[1].transAxes)
    axes[1].axis('off')

@@

# Display uncertainty information if computed
if 'velocity_pyrovelocity_uncertainty' in adata_with_velocity.var.columns:
    print("Velocity uncertainty statistics:")
    uncertainty = adata_with_velocity.var['velocity_pyrovelocity_uncertainty']
    print(f"Mean uncertainty: {uncertainty.mean():.4f}")
    print(f"Std uncertainty: {uncertainty.std():.4f}")
    print(f"Min uncertainty: {uncertainty.min():.4f}")
    print(f"Max uncertainty: {uncertainty.max():.4f}")
```

>>>>

```python
import torch
import pyro
import numpy as np
import matplotlib.pyplot as plt
import scanpy as sc
import scvelo as scv
from pyrovelocity.utils import print_anndata, anndata_string, print_string_diff
from pyrovelocity.models.modular import PyroVelocityModel
from pyrovelocity.models.modular.factory import create_legacy_model2
from pyrovelocity.io.datasets import bifurcation_14

@@

# Create a legacy model (add_offset=True) that replicates the original PyroVelocity implementation
model = create_legacy_model2()
print(model)

@@

# Generate posterior samples and compute velocity in one step
# The new unified get_velocity method computes both velocity and uncertainty
adata_with_velocity = model.get_velocity(
    adata=adata,
    num_samples=100,  # Use more samples for better uncertainty estimates
    basis='umap',
    compute_uncertainty=True,
    uncertainty_method='std',
    model_name='legacy2'
)

@@

# Train the custom legacy model
custom_model.train(
    adata=adata_with_velocity,  # Use the data with embeddings
    max_epochs=50,
    learning_rate=0.01,
    use_gpu=False
)

# Compute velocity with the custom legacy model
custom_adata = custom_model.get_velocity(
    adata=adata_with_velocity,
    num_samples=100,
    basis='umap',
    compute_uncertainty=True,
    model_name='custom_legacy2'
)
```

@@

```python
from pyrovelocity.models.modular.factory import (
    create_model_from_config,
    create_legacy_model1,
    create_legacy_model2
)
from pyrovelocity.models.modular.config import ModelConfig

# The primary model we use is create_legacy_model2 (add_offset=True)
# This replicates the original PyroVelocity implementation most closely
primary_model = create_legacy_model2()

# Alternative: create_legacy_model1 (add_offset=False)
alternative_model = create_legacy_model1()

# You can also create models from configurations
config = ModelConfig.standard()
config_model = create_model_from_config(config)

# Train and use the primary legacy model
primary_model.train(adata=adata_with_velocity, max_epochs=50)

# The new unified get_velocity method returns the mutated AnnData object
adata_primary = primary_model.get_velocity(
    adata=adata_with_velocity,
    num_samples=100,
    basis='umap',
    compute_uncertainty=True,
    model_name='primary_legacy2'
)

print("Available velocity layers:")
for layer in adata_primary.layers.keys():
    if 'velocity' in layer:
        print(f"- {layer}")
```

@@

# Plot 1: Primary legacy model (create_legacy_model2)
scv.pl.velocity_embedding_grid(
    adata_with_velocity,
    basis='umap',
    color='leiden',
    title='PyroVelocity Legacy Model 2 (add_offset=True)',
    vkey='velocity_legacy2',  # Use the velocity key added by get_velocity
    s=15,
    alpha=1,
    linewidth=0.5,
    ax=axes[0],
    show=False,
    legend_loc='right margin',
    density=0.4,
    scale=0.2,
    arrow_size=2,
    arrow_length=2,
    arrow_color='black',
)

# Plot 2: Alternative model comparison (if computed)
if 'velocity_primary_legacy2' in adata_primary.layers:
    scv.pl.velocity_embedding_grid(
        adata_primary,
        basis='umap',
        color='leiden',
        title='PyroVelocity Primary Legacy Model',
        vkey='velocity_primary_legacy2',
        s=15,
        alpha=1,
        linewidth=0.5,
        ax=axes[1],
        show=False,
        legend_loc='right margin',
        density=0.4,
        scale=0.2,
        arrow_size=2,
        arrow_length=2,
        arrow_color='black',
    )
else:
    axes[1].text(0.5, 0.5, 'Alternative model\nnot computed',
                ha='center', va='center', transform=axes[1].transAxes)
    axes[1].axis('off')

@@

# Display uncertainty information if computed
if 'velocity_legacy2_uncertainty' in adata_with_velocity.var.columns:
    print("Velocity uncertainty statistics:")
    uncertainty = adata_with_velocity.var['velocity_legacy2_uncertainty']
    print(f"Mean uncertainty: {uncertainty.mean():.4f}")
    print(f"Std uncertainty: {uncertainty.std():.4f}")
    print(f"Min uncertainty: {uncertainty.min():.4f}")
    print(f"Max uncertainty: {uncertainty.max():.4f}")
```
